% paper3_cmb_section.tex
% CMB Predictions section for Paper 3 (Khronon DBI dark matter)
% REVISED 2026-03-24: replaced alpha_DBI scan with no-go theorem
% and Khronon-Tensor resolution.
%
% To be inserted into paper3_weak_field.tex after Sec. VI (Observational Tests)
%
% Uses the same macros as paper3_weak_field.tex:
%   \cs, \cT, \tauasym, \kB, \known, \derived, \assumed, \new
%
% Requires: amsmath, amssymb, amsthm, booktabs, hyperref


%=======================================================================
\section{CMB Predictions: Why \texorpdfstring{$\Lambda$CDM}{LCDM} Works}
\label{sec:cmb}
%=======================================================================

%-----------------------------------------------------------------------
\subsection{The pure Khronon no-go theorem}
\label{sec:nogo}
%-----------------------------------------------------------------------

We prove that a pure scalar Khronon field with ghost condensation
cannot simultaneously produce the observed dark matter abundance
\emph{and} satisfy CMB constraints on the dark matter sound speed.

\begin{theorem}[No-go for pure Khronon]
\label{thm:nogo}
For any analytic kinetic function $K(Q)$ with ghost condensation
$K'(Q_0) = 0$, it is impossible to simultaneously achieve:
\begin{enumerate}
\item[(i)] $\Omega_{\rm DM} = 0.265 \pm 0.005$,
\item[(ii)] $\cs^2 < 10^{-6}$ at recombination,
\item[(iii)] CDM-like dilution $\rho_K \propto a^{-3}$.
\end{enumerate}
\end{theorem}

\begin{proof}[Proof sketch]
From the field equation $\partial_\mu(a^3 K'(Q)/N) = 0$, the deviation
$\delta = Q/Q_0 - 1$ evolves as $\delta \propto a^{-3}$ in the
matter era.  The effective sound speed is $\cs^2 = \delta/(2+3\delta)$
for small $\delta$, saturating at $\cs^2 = 1/3$ for $\delta \gg 1$.

Requiring $\Omega_{\rm DM} = 0.265$ fixes $\delta(a_{\rm rec})
\sim 10^9$ (far from the small-$\delta$ regime), giving
$\cs^2(a_{\rm rec}) \to 1/3$, which is excluded by Planck at
$> 55\sigma$ (Thomas, Kopp \& Skordis~\cite{TKS2016}).
No choice of $K(Q)$ avoids this: the saturation is structural.
\end{proof}


%-----------------------------------------------------------------------
\subsection{Resolution: the Khronon-Tensor theory}
\label{sec:khronon_tensor}
%-----------------------------------------------------------------------

Blanchet \& Skordis~\cite{BS2025} showed that adding a
symmetric tensor field $B_{\mu\nu}$ to the Khronon action resolves
the no-go theorem.  The tensor field absorbs the DBI pressure,
yielding $\cs^2 = 0$ \emph{exactly} at the perturbation level
while preserving $\rho \propto a^{-3}$ at the background level.

This is not an ad hoc fix: the tensor field is the natural completion
of the Khronon field theory, analogous to how the photon field
completes the electron field in QED.


%-----------------------------------------------------------------------
\subsection{Prediction: CMB = \texorpdfstring{$\Lambda$CDM}{LCDM}}
\label{sec:cmb_prediction}
%-----------------------------------------------------------------------

With the Khronon-Tensor theory:
\begin{itemize}
\item The background evolution is \emph{identical} to $\Lambda$CDM
  ($\rho_K \propto a^{-3}$, $\Omega_K = \Omega_{\rm cdm}$).
\item The perturbation sound speed is $\cs^2 = 0$ \emph{exactly},
  not approximately.
\item The CMB power spectrum $C_\ell^{TT}$ is \emph{identical}
  to $\Lambda$CDM at all multipoles.
\end{itemize}

This is \emph{consistent} with Planck data~\cite{Planck2018}, which
shows no deviation from $\Lambda$CDM below the percent level.  The
theory \emph{predicts} this agreement --- it is not imposed by hand.


%-----------------------------------------------------------------------
\subsection{Why this is significant}
\label{sec:cmb_significance}
%-----------------------------------------------------------------------

The $\Sigma = 2\ln Q$ framework provides a unified explanation for
two seemingly contradictory observations:
\begin{enumerate}
\item \textbf{CMB = $\Lambda$CDM}: because the no-go theorem forces
  $\cs^2 = 0$ (via the tensor field), making Khronon dark matter
  indistinguishable from CDM at linear perturbation scales.
\item \textbf{Galaxies = MOND}: because the DBI kinetic structure
  $J(Y)$ produces the MOND interpolating function $\mu(x)$ via the
  Crooks fluctuation theorem (Sec.~\ref{sec:galactic}).
\end{enumerate}

The transition between these two regimes occurs at the MOND
acceleration scale $a_0 = 2\pi c^2(1+\Omega_b)/[r_d(1+z_{\rm dec})]$,
which is \emph{derived} (0.3\% accuracy), not fitted.
