% paper3_weak_field.tex
% Dark Matter Phenomenology from Khronon Condensation Without Dark Matter Particles
% of the tau Framework
% For submission to Physical Review D
% Author: Sheng-Kai Huang
% Compile: pdflatex paper3_weak_field.tex && bibtex paper3_weak_field && pdflatex paper3_weak_field && pdflatex paper3_weak_field

\documentclass[prd,twocolumn,superscriptaddress,longbibliography,floatfix]{revtex4-2}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{xcolor}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem*{proposition}{Proposition}

% Convenience macros
\newcommand{\kB}{k_{\mathrm{B}}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\Msun}{M_\odot}
\newcommand{\tauasym}{\tau}
\newcommand{\cs}{c_{\mathrm{s}}}
\newcommand{\cT}{c_{\mathrm{T}}}

% Epistemic tags
\newcommand{\known}{\textsc{[known]}}
\newcommand{\derived}{\textsc{[derived]}}
\newcommand{\assumed}{\textsc{[assumed]}}
\newcommand{\new}{\textsc{[new synthesis]}}

\begin{document}

\title{Dark Matter Phenomenology from Khronon Condensation\\
Without Dark Matter Particles}

\author{Sheng-Kai Huang}
\email{akai@fawstudio.com}
\affiliation{Independent Researcher}

\date{\today}

\begin{abstract}
The temporal asymmetry parameter $\tauasym = 1 - F$, introduced in
Paper~1 as the Petz recovery failure, acquires physical realization
through the Blanchet--Skordis (BS) Khronon action
$S = \int\!\sqrt{-g}\,[R - 2J(Y) + 2K(Q)]\,d^4x$.
This action contains two structurally independent sectors: the
\emph{cosmological sector} $K(Q) = \mu^2(Q-1)^2$, fixed by ghost
condensation with a single parameter $\mu_0 = H_0/c$ from
dimensional analysis, which reproduces the observed dark matter
abundance $\Omega_{\mathrm{DM}} = \delta_0(2{+}\delta_0)/3 = 0.265$
with CDM-like perturbations ($\cs^2 = 0$) \emph{without invoking
any dark matter particle}; and the
\emph{galactic sector} $J(Y)$, a phenomenological free function
constrained only at its asymptotic limits, which reproduces the
Baryonic Tully--Fisher Relation and the Radial Acceleration Relation
with one empirical parameter $a_0 \approx 1.2 \times 10^{-10}\;\mathrm{m/s}^2$.
We prove a no-go theorem establishing that pure scalar Khronon dark
matter is incompatible with CMB observations, necessitating the
Blanchet--Skordis tensor field. The resulting Khronon-Tensor theory
predicts a CMB power spectrum \emph{identical} to $\Lambda$CDM ---
consistent with Planck data --- while producing MOND dynamics at
galaxy scales.
We present a systematic confrontation of this two-sector framework
with 11~low-redshift observational tests (10~passed, 1~marginal),
show that the Bullet Cluster mass-gas offset is naturally explained
by the $\cs^2 = 0$ collisionless Khronon condensate, compare its
parameter economy with $\Lambda$CDM and AeST, and identify
falsifiable predictions.
We are explicit about what is derived, what is assumed, and what
remains open---in particular, the connection between $a_0$ and
$\mu_0$ is suggestive but unproven, and the running
$\mu(k) = k$ is a well-motivated assumption, not a derivation.
\end{abstract}

\maketitle


%=======================================================================
\section{Introduction}
\label{sec:intro}
%=======================================================================

The dark matter problem encompasses two distinct observational
domains: (i)~cosmological---the matter density inferred from
the cosmic microwave background (CMB), baryon acoustic oscillations
(BAO), and large-scale structure; and (ii)~galactic---the mass
discrepancy observed in galaxy rotation curves, the Baryonic
Tully--Fisher Relation (BTFR), and the Radial Acceleration
Relation (RAR).  The standard $\Lambda$CDM model addresses both
with a single ingredient---cold dark matter (CDM) particles---but
despite four decades of searches, no particle has been directly
detected~\cite{Schumann2019,XENON2023,LZ2024}.

Modified gravity approaches (MOND~\cite{Milgrom1983},
TeVeS~\cite{Bekenstein2004}, AeST~\cite{SZ2021prl,SZ2021extended})
excel at galactic phenomenology but have historically struggled
with cosmology.  The breakthrough of Skordis and
Zl\'{o}snik~\cite{SZ2021prl} demonstrated that a relativistic
MOND theory \emph{can} reproduce both the CMB power spectrum and
the matter power spectrum, provided the theory contains a scalar
field whose perturbations mimic CDM at early times ($\cs^2 = 0$).

In Paper~1~\cite{Paper1}, we introduced the temporal asymmetry
parameter $\tauasym = 1 - F(\rho,\,\mathcal{R}_{\sigma,\mathcal{N}}
(\mathcal{N}(\rho)))$, which quantifies the arrow of time as Petz
recovery failure, with the universal bound
$\tauasym \leq 1 - \exp(-\Sigma/2)$.  In Paper~2~\cite{Paper2},
we showed that in strong gravitational fields
$\Sigma_{\mathrm{grav}} = -\ln(-g_{00})$, connecting the Petz
recovery fidelity to the gravitational channel transmissivity
$\eta = -g_{00}$.

\emph{Three long-standing problems of MOND.}---Since Milgrom's
original proposal~\cite{Milgrom1983}, three fundamental questions
have remained unanswered:
\begin{enumerate}
\item \emph{Why $a_0 \approx cH_0$?}---The MOND acceleration
  scale $a_0 \approx 1.2 \times 10^{-10}\;\mathrm{m/s}^2$ is
  numerically close to $cH_0/(2\pi)$, a cosmological quantity.
  No MOND theory has explained this ``Milgrom coincidence.''
\item \emph{Why does the CMB look like $\Lambda$CDM?}---MOND
  modifies gravity at all scales, yet the CMB acoustic peaks
  are exquisitely fit by $\Lambda$CDM with pressureless cold dark
  matter.  Previous relativistic completions (TeVeS, AeST) require
  additional fields and parameters to reconcile this.
\item \emph{What determines $\mu(x)$?}---The MOND interpolating
  function $\mu(x)$ has been a free function for forty years,
  with multiple proposals but no theoretical selection principle.
\end{enumerate}
The $\Sigma = 2\ln Q$ framework addresses these three questions
through distinct mechanisms: $\mu_0 = H_0/c$ is the unique scale
without Planck physics, providing a \emph{theoretical} answer to
question~1; the no-go theorem (Theorem~\ref{thm:nogo}) forces
$\cs^2 = 0$ via the tensor field, providing a \emph{mathematical}
answer to question~2; and the Petz recovery bound constrains the
interpolating function to $\mu(x) = 1 - e^{-\sqrt{x}}$, providing
a \emph{testable candidate} for question~3---not a derivation, since
the deep-MOND asymptotic behavior (BTFR) is used as a boundary
condition, but the transition-region shape is a genuine prediction
verifiable against SPARC data (Sec.~\ref{sec:sparc_test}).

This paper addresses the \emph{weak-field} phenomenology of the
same framework.  We adopt the Blanchet--Skordis (BS) Khronon
action~\cite{BS2024} as the concrete realization of the $\tauasym$
framework at galactic and cosmological scales.  Our goals are:
\begin{enumerate}
\item To present the BS Khronon theory with honest parameter
  counting and epistemic transparency;
\item To show that $K(Q) = \mu^2(Q-1)^2$ with $\mu_0 = H_0/c$
  reproduces cosmological dark matter naturally;
\item To show that $J(Y)$ reproduces MOND galactic phenomenology;
\item To systematically confront the theory with low-redshift
  observations;
\item To identify what is derived, what is assumed, and what
  remains open.
\end{enumerate}

\emph{What this paper does not claim.}---We do not claim to
``solve'' the dark matter problem.  We provide a new interpretation
in which all dark matter phenomenology---flat rotation curves,
the CMB acoustic peaks, Bullet Cluster offsets, and the
matter power spectrum---emerges from the ghost condensation of
a Khronon field, \emph{without any new particle species}.
If this framework is correct, direct detection experiments
(LZ, XENONnT, PandaX) and collider searches should yield null
results for dark matter particles.
The theory contains a free function $J(Y)$ (the MOND interpolating
function), and two key assumptions (running $\mu$, the
$a_0$--$\mu_0$ coincidence) remain unproven.


%=======================================================================
\section{The Blanchet--Skordis Khronon Action}
\label{sec:action}
%=======================================================================

\subsection{Action and field content}

\known\  The BS Khronon theory~\cite{BS2024,BS2025} describes
gravity coupled to a single scalar field $\phi$ (the ``Khronon'')
whose level surfaces define spatial hypersurfaces.  The action is
\begin{equation}
S = \frac{c^3}{16\pi G}\int\!d^4x\,\sqrt{-g}\,
\big[R - 2J(Y) + 2K(Q)\big] + S_{\mathrm{m}},
\label{eq:action}
\end{equation}
where $S_{\mathrm{m}}$ is the matter action (universally coupled
to $g_{\mu\nu}$ only), and:
\begin{itemize}
\item The \emph{unit normal} to the $\phi = \mathrm{const}$
  hypersurfaces is
  $n_\mu = -(c/Q)\,\nabla_\mu\phi$, with
  $Q = c\sqrt{-g^{\mu\nu}\nabla_\mu\phi\,\nabla_\nu\phi}$.
\item The \emph{acceleration} of the foliation is
  $A_\mu = c^2\,n^\nu\nabla_\nu n_\mu$, with
  $Y = A_\mu A^\mu / c^4$.
\item $K(Q)$ controls the \textbf{cosmological sector}:
  the energy density and equation of state of the Khronon field
  on the FRW background.
\item $J(Y)$ controls the \textbf{galactic sector}:
  the quasi-static, nonrelativistic limit that reproduces
  MOND phenomenology.
\end{itemize}

\known\  The Khronon field is \emph{hypersurface-orthogonal by
construction} ($n_\mu$ is a gradient), which eliminates the twist
tensor $\omega_{\mu\nu} = 0$ and merges the Einstein--Aether
parameters $c_1$ and $c_3$ into the single combination
$c_{13} = c_1 + c_3$~\cite{Jacobson2010,BPS2010,Horava2009}.

\subsection{Why $c_{13} = 0$: action choice, not geometry}
\label{sec:c13}

\assumed\  The BS action~\eqref{eq:action} does \emph{not} contain
the extrinsic curvature coupling $K_{\mu\nu}K^{\mu\nu}$ (where
$K_{\mu\nu}$ is the extrinsic curvature of the $\phi = \mathrm{const}$
hypersurfaces).  In the Einstein--Aether parametrization, this
coupling generates $c_{13} \neq 0$.  Its absence yields
$c_{13} = 0$ \emph{by construction}~\cite{BS2024}, giving the
gravitational wave speed
\begin{equation}
\cT^2 = \frac{1}{1 - c_{13}} = 1,
\quad\Longrightarrow\quad \cT = c \;\;\text{(exactly)}.
\label{eq:cT}
\end{equation}
This is a \emph{choice of action}, not a consequence of hypersurface
orthogonality alone (which only merges $c_1, c_3$ into $c_{13}$ but
does not set it to zero).  The choice is strongly motivated by:
\begin{enumerate}
\item GW170817~\cite{GW170817}: $|c_{\mathrm{T}}/c - 1| < 10^{-15}$;
\item Theoretical minimality (fewer terms in the action);
\item Even if a $K_{\mu\nu}K^{\mu\nu}$ term were added, the tiny
  coupling $\mu_0 = H_0/c$ would make $c_{13} \sim 10^{-30}$ at
  solar system scales~\cite{OostMW2018}.
\end{enumerate}

\subsection{Degrees of freedom and stability}
\label{sec:stability}

\known\  The hypersurface-orthogonal Khronon propagates 3~degrees
of freedom~\cite{BPS2010}: 2~tensor (spin-2 graviton) + 1~scalar
(Khronon mode).  Vector modes are eliminated by the constraint
$\omega_{\mu\nu} = 0$.

For the kinetic function $K(Q)$ with a minimum at $Q = Q_0$,
the scalar sector is free of all standard instabilities~\cite{AHCLM2004,BS2024}:
\begin{itemize}
\item \textbf{No Ostrogradski ghost}: the field equation is
  second-order in $\phi$ (BPS~\cite{BPS2010}, Hamiltonian analysis);
\item \textbf{No wrong-sign kinetic term}: $K''(Q_0) > 0$ ensures
  positive-definite kinetic energy;
\item \textbf{No gradient instability}: $\cs^2 = K'(Q_0)/
  [K'(Q_0) + 2Q_0\,K''(Q_0)] = 0$ (marginal, cured by
  ghost condensation $k^4$ terms~\cite{AHCLM2004});
\item \textbf{No tachyon}: $Q_0$ is a global minimum of $K$;
\item \textbf{$\cT = c$ exactly}: $c_{13} = 0$ by construction;
\item \textbf{Strong coupling at $\Lambda_3 \sim (\mu_0^2 M_{\mathrm{Pl}})^{1/3}
  \sim 10^{-13}$~eV}: far above all cosmological and astrophysical
  scales ($H_0 \sim 10^{-33}$~eV).
\end{itemize}
The full 8-criterion stability analysis is presented in the
Supplemental Material~\cite{stability_SM}.

\subsection{Honest parameter counting}
\label{sec:params}

The theory contains:
\begin{itemize}
\item 6 standard cosmological parameters
  ($H_0$, $\Omega_{\mathrm{b}}h^2$, $n_s$, $A_s$, $\tau_{\mathrm{reion}}$,
  $\Sigma m_\nu$);
\item 1 cosmological parameter: $\mu$ (Khronon mass, or
  equivalently $\Omega_{\mathrm{DM}}h^2$ as an initial condition);
\item 1 galactic parameter: $a_0$ (MOND acceleration scale);
\item 1 free function: $J(Y)$ (MOND interpolating function,
  constrained only at asymptotic limits).
\end{itemize}

\begin{table}[b]
\caption{\label{tab:params}Parameter counting comparison.
``func'' denotes a free function.
$\Omega_{\mathrm{DM}}h^2$ is an initial condition (fitted, not
predicted) in all theories, including $\Lambda$CDM.}
\begin{ruledtabular}
\begin{tabular}{lccc}
\textbf{Theory} & \textbf{Cosmo} & \textbf{Theory-specific}
  & \textbf{Free func.} \\
\colrule
$\Lambda$CDM & 6 & $\Omega_{\mathrm{DM}}h^2$ & 0 \\
MOND (non-rel.) & --- & $a_0$ & $\mu(x)$ \\
AeST~\cite{SZ2021prl} & 6 & $K_B, K_2, \lambda_s, Q_0, a_0$ & $F(Y,Q)$ \\
BS Khronon & 6 & $\mu_0, a_0, \Omega_{\mathrm{DM}}h^2$ & $J(Y)$ \\
BS Khronon + $\tauasym$ & 6 & $a_0, \Omega_{\mathrm{DM}}h^2$ & $J(Y)$ \\
\end{tabular}
\end{ruledtabular}
\end{table}

Table~\ref{tab:params} compares parameter counts across theories.
The $\tauasym$ framework fixes $\mu_0 = H_0/c$ via dimensional
analysis (Sec.~\ref{sec:mu0}), reducing the theory-specific
parameter count by one relative to plain BS Khronon.  The free
function $J(Y)$ is inherited from the MOND tradition dating to
Bekenstein and Milgrom~\cite{BekensteinMilgrom1984}, and has been
free for forty years.  The Crooks interpolation function
$\mu(x) = 1 - e^{-\sqrt{x}}$ (Sec.~\ref{sec:crooks}) could in
principle eliminate this freedom if validated against SPARC data.


%=======================================================================
\section{Cosmological Sector: $K(Q)$}
\label{sec:cosmo}
%=======================================================================

\subsection{Ghost condensation fixes $K(Q)$}
\label{sec:KQ}

\derived\  Any scalar field that spontaneously breaks Lorentz
invariance by acquiring a timelike gradient
$\langle\partial_\mu\phi\rangle \neq 0$ sits in the ``ghost
condensation'' regime~\cite{AHCLM2004}.  The leading-order
effective action around the condensate $Q = 1$ is
\begin{equation}
K(Q) = \mu^2\,(Q - 1)^2,
\label{eq:KQ}
\end{equation}
as established by Arkani-Hamed et al.~\cite{AHCLM2004} and
Scherrer~\cite{Scherrer2004}.  The quadratic form is not
arbitrary: it is the unique leading-order expansion around
a kinetic minimum with $K'(1) = 0$ and $K''(1) = 2\mu^2 > 0$,
ensuring stability and $\cs^2 = 0$.  Petz optimality provides
an independent argument: the maximally recoverable channel
configuration corresponds to a quadratic kinetic function.

Higher-order terms $K(Q) = \mu^2(Q-1)^2 + \alpha_3(Q-1)^3 + \cdots$
are possible.  The DBI completion~\cite{BS2024}
$K_{\mathrm{DBI}} = \mu^2\lambda_D^2[\sqrt{1 + (Q-1)^2/\lambda_D^2} - 1]$
provides a natural UV regularization, introducing at most one
additional parameter $\lambda_D$.  We focus on the leading
quadratic form throughout.

\subsection{$\mu_0 = H_0/c$: dimensional analysis}
\label{sec:mu0}

\assumed\  The coupling $\mu$ has dimensions $[\mathrm{m}^{-1}]$.
The available fundamental constants are
$\{H_0, c, \hbar, G, \kB\}$.  The general combination with
these dimensions is~\cite{muSynthesis}
\begin{equation}
\mu = \frac{H_0}{c}\,\left(t_{\mathrm{P}} H_0\right)^{2d},
\label{eq:mu_general}
\end{equation}
where $t_{\mathrm{P}} = \sqrt{\hbar G/c^5}$ is the Planck time
and $t_{\mathrm{P}} H_0 = 1.18 \times 10^{-61}$.  For \emph{any}
$d \neq 0$, the correction factor is $10^{-122|d|}$, destroying
the result.  Therefore
\begin{equation}
\boxed{\mu_0 = \frac{H_0}{c} = 7.28 \times 10^{-27}\;\mathrm{m}^{-1}}
\label{eq:mu0}
\end{equation}
is the \emph{unique} selection without Planck-scale physics.  This
is a naturality argument, not a derivation from field equations.
Three equivalent physical interpretations hold:
\begin{enumerate}
\item Khronon Compton wavelength $= c/H_0 =$ Hubble radius;
\item $\mu_0 = 2\pi\kB T_{\mathrm{dS}}/(\hbar c)$: inverse
  de~Sitter thermal wavelength~\cite{GibbonsHawking1977};
\item $\mu_0 = 2\pi a_0/c^2$: connects to the MOND scale
  (Sec.~\ref{sec:a0_mu}).
\end{enumerate}

\subsection{Cosmological background: FRW solution}
\label{sec:FRW}

\known\  On the FRW background, the Khronon field equation
gives~\cite{BS2024}
\begin{equation}
K'(Q_0) = \frac{I_0}{a^3 Q_0},
\label{eq:khronon_FRW}
\end{equation}
where $I_0$ is an integration constant (the initial Khronon
momentum).  This conservation law holds even when $\mu$
varies in time, since the Lagrangian has no explicit
$\phi$-dependence ($\partial\mathcal{L}/\partial\phi = 0$).
Substituting $K'(Q_0) = 2\mu^2\delta$ gives
\begin{equation}
Q_0(a) = 1 + \delta(a), \qquad
\delta(a) = \frac{I_0}{2\mu^2(a)\,a^3},
\label{eq:delta}
\end{equation}
where $\mu(a)$ is the (possibly running) coupling.

A crucial self-consistency: regardless of how $\mu$ runs,
the combination $\mu^2\delta = I_0/(2a^3)$ depends only on
$a$.  Therefore the Khronon energy
density~\eqref{eq:rhoK} satisfies
$\rho_K \propto (2+\delta)/a^3$, which is CDM-like
($\rho_K \propto a^{-3}$) whenever $(2+\delta)$ varies
slowly---i.e., whenever $\delta \lesssim O(1)$.

For constant $\mu = \mu_0$:
$\delta \propto a^{-3}$, giving $\delta(z=1100)
\sim 10^9 \gg 1$---the Khronon exits the ghost
condensation regime and the energy density scales as
$\rho_K \propto a^{-6}$ (stiff matter), incompatible
with CMB observations.  This is the true
\textbf{CMB Catch-22} (Sec.~\ref{sec:running_mu}).
Running $\mu_{\mathrm{bg}} = H(a)/c$ keeps
$\delta \lesssim O(1)$ at all epochs
(Sec.~\ref{sec:running_mu}), maintaining the ghost
condensation regime and CDM-like dilution.

The Khronon energy density follows from the k-essence
stress-energy $\rho = QK' - K$~\cite{BS2024}:
\begin{equation}
\rho_K = \frac{c^2\mu^2\,\delta\,(2+\delta)}{8\pi G},
\label{eq:rhoK}
\end{equation}
where the factor $(2+\delta)$ arises from
$QK' - K = (1{+}\delta)\cdot 2\mu^2\delta - \mu^2\delta^2
= \mu^2\delta(2{+}\delta)$.
For $\delta \ll 1$, $\rho_K \approx c^2\mu^2\delta/({4\pi G})$;
for $\delta \gg 1$, $\rho_K \sim c^2\mu^2\delta^2/(8\pi G)$.
Since $\mu^2\delta = I_0/(2a^3)$, one has
$\rho_K = c^2 I_0\,(2+\delta)/(16\pi G\,a^3)$:
the energy density dilutes as $a^{-3}$ (CDM-like)
with $O(1)$ corrections encoded in $(2+\delta(a))$.

The equation of state follows from $w = K/\rho_K
= \mu^2\delta^2/[\mu^2\delta(2{+}\delta)]$~\cite{BS2024}:
\begin{equation}
w = \frac{\delta}{2+\delta}.
\label{eq:w_eos}
\end{equation}
\emph{Importantly}, $w$ does \emph{not} map
directly to the GDM equation of state
$w_{\mathrm{GDM}}$ constrained by
Thomas, Kopp \& Skordis~\cite{TKS2016}.
The actual instantaneous dilution rate is
$d\ln\rho_K/d\ln a = -3 - d\ln(2{+}\delta)/d\ln a$,
which equals $-3$ whenever $\delta$ varies slowly.

\begin{theorem}[Khronon Information Conservation]
\label{thm:conservation}
For the Blanchet--Skordis action with $K(Q) = \mu^2(Q-1)^2$ on
FRW background, the scalar field equation implies
\begin{equation}
a^3\,Q_0\,K'(Q_0) \;=\; I_0 \;=\; \mathrm{const},
\label{eq:conservation}
\end{equation}
where $I_0$ is the conserved Khronon momentum.
This holds for arbitrary time-dependent $\mu(t)$.
\end{theorem}
\begin{proof}[Proof sketch]
The Lagrangian density $\mathcal{L}(\partial\phi,\partial^2\phi)$
has no explicit $\phi$-dependence
($\partial\mathcal{L}/\partial\phi = 0$).  By the Euler--Lagrange
equation, the canonical current
$J^a = \partial\mathcal{L}/\partial(\partial_a\phi)$ satisfies
$\partial_a(\sqrt{-g}\,J^a) = 0$.  On FRW with $\phi = t + \delta\phi(t)$,
the spatial components vanish by isotropy, leaving the temporal
conservation law~\eqref{eq:conservation}.
\end{proof}
\noindent\textit{Consequences:}
\begin{enumerate}
\item \textbf{CDM-like dilution identity}:
  $\mu^2\delta = I_0/(2a^3)$---the product is independent of
  how $\mu$ runs.
\item \textbf{Running $\mu$ stability}: with $\mu = H(a)/c$,
  $\delta(a) = I_0 c^2/(2H^2 a^3)$ stays $O(1)$ at all epochs.
\item \textbf{Automatic $a^{-3}$ scaling}:
  $\rho_K = c^2 I_0\,(2{+}\delta)/(16\pi G\,a^3)$ dilutes as
  dust whenever $(2{+}\delta) \approx \mathrm{const}$.
\end{enumerate}

\noindent\textit{Remark (Unified entropy production).}---On
static backgrounds $Q = 1/\sqrt{-g_{00}}$, giving
$\Sigma_{\mathrm{grav}} = -\ln(-g_{00}) = 2\ln Q$
(Paper~2~\cite{Paper2}).  On FRW backgrounds $Q_0 = 1 + \delta$,
giving $\Sigma_{\mathrm{cosmo}} = 2\ln(1+\delta)$.
Both are instances of the general result
$\Sigma = 2\ln Q$, where $Q$ is the inverse Khronon lapse.
This identity holds exactly on stationary and FRW backgrounds,
and to $O(\varepsilon_{\mathrm{NA}}^2)$ on any adiabatic
background, where
$\varepsilon_{\mathrm{NA}} = |d\ln Q/d\tau|/(2\omega)$
is the non-adiabaticity parameter of the probe mode.
The correction is always non-negative (particle creation adds
entropy), so $\Sigma \geq 2\ln Q$ in general.

\subsection{Running $\mu$: assumption and motivation}
\label{sec:running_mu}

\assumed\  With constant $\mu = \mu_0 = H_0/c$,
$\delta(a) = \delta_0/a^3$ grows as $a^{-3}$, reaching
$\delta \sim 10^9$ at $z = 1100$.  In this regime
the Khronon exits ghost condensation:
$(2{+}\delta) \sim \delta$, so
$\rho_K \propto \delta^2/a^3 \propto a^{-9}$
(stiff matter, $w \to 1$).  This is the
\textbf{CMB Catch-22}: constant $\mu_0 = H_0/c$ drives
$\delta \gg 1$ at early times, destroying CDM-like behavior.

We \emph{assume} the running
\begin{equation}
\mu_{\mathrm{bg}}(a) = \frac{H(a)}{c},
\label{eq:mu_running}
\end{equation}
motivated by three convergent arguments:
\begin{enumerate}
\item \textbf{DPI + extensivity}: the anomalous dimension
  $\eta_\mu = d - 3 = 1$ in $d = 4$ follows from the transition
  from area-law to volume-law entanglement entropy at the
  de~Sitter scale, combined with the data processing
  inequality~\cite{Verlinde2016,Jacobson2016,CasiniHuerta2017,DorauMuch2025};
\item \textbf{Dimensional transmutation}: $\mu(k) = k$ is the
  unique power-law running independent of the boundary condition
  $H_0$; Kumar~\cite{Kumar2025} obtains a logarithmic correction
  to Newton's potential from marginal IR running in QFT,
  providing independent support;
\item \textbf{Modular flow}: if the Khronon is the modular flow
  direction, the modular temperature $T_{\mathrm{mod}} = k/\pi$
  gives $\mu(k) \sim k$.
\end{enumerate}
\emph{A rigorous derivation from the microscopic theory remains
an open problem.}

With~\eqref{eq:mu_running},
$\delta(a) = I_0 c^2/(2H^2(a)\,a^3)$, which gives:
\begin{itemize}
\item \textbf{Radiation era} ($H^2 \propto a^{-4}$):
  $\delta \propto a \to 0$ at early times;
\item \textbf{Matter era} ($H^2 \propto a^{-3}$):
  $\delta \approx \delta_0/\Omega_m \sim O(1)$;
\item \textbf{Today}: $\delta_0 = 0.340$,
  $w_0 = \delta_0/(2{+}\delta_0) = 0.145$.
\end{itemize}
The running keeps $\delta \lesssim O(1)$ at all epochs,
maintaining the ghost condensation regime.  Since
$\mu^2\delta = I_0/(2a^3)$ is independent of the running,
the Khronon energy density $\rho_K \propto (2{+}\delta)/a^3$
remains CDM-like, with $O(1)$ corrections during the
matter era that require a full Boltzmann solver to quantify.
The perturbation sound speed $\cs^2 = 0$ is unaffected.

\subsection{Perturbations: $\cs^2 = 0$ (CDM-like)}
\label{sec:perturbations}

\derived\  The scalar sound speed is~\cite{BS2024}
\begin{equation}
\cs^2 = \frac{K'(Q_0)}{K'(Q_0) + 2Q_0\,K''(Q_0)}
\;\xrightarrow{K'(1)=0}\; 0.
\label{eq:cs2}
\end{equation}
This is not fine-tuned: it follows from $K(Q)$ having a
\emph{minimum} at $Q_0 = 1$ (ghost condensation).  The
$\cs^2 = 0$ condition is radiatively stable, with quantum
corrections $\delta\cs^2 \sim (\mu/M_{\mathrm{Pl}})^2
\sim 10^{-122}$~\cite{AHCLM2004}.

\emph{Consequence}: Khronon perturbations grow identically to
CDM at all cosmologically relevant scales.  The Jeans length
$\lambda_J = 0$ (no pressure support), so structure formation
proceeds as in $\Lambda$CDM.  In particular, the CMB acoustic
peak structure, matter power spectrum $P(k)$, and BAO features
are reproduced without modification.

\emph{Physical meaning of $\cs^2 = 0$.}---The vanishing sound
speed means Khronon perturbations do \emph{not} propagate: they
grow in place, exactly like CDM density perturbations.
By Theorem~\ref{thm:conservation}, $K'(Q_0) = I_0/(a^3 Q_0)
\to 0$ as the condensate approaches $Q_0 = 1$; since the
numerator in~\eqref{eq:cs2} is $K'(Q_0)$ while the denominator
retains $2Q_0 K''(Q_0) = 4\mu^2 > 0$, the ratio vanishes
identically at the ghost condensation point.
This resolves the apparent tension with the GDM constraints of
Thomas, Kopp \& Skordis~\cite{TKS2016}: the BS effective equation
of state $w = \delta/(2{+}\delta)$~\eqref{eq:w_eos} does \emph{not}
equal the GDM sound speed $c_{\mathrm{s,GDM}}^2$.  The two quantities
are structurally independent---$w$ governs the background
dilution rate while $\cs^2$ is set by the kinetic structure of
$K(Q)$.  Since $\cs^2 = 0$ exactly, the TKS bound
$\cs^2 < 3.21 \times 10^{-6}$ (99.7\%~CL) is trivially
satisfied.

\subsection{Cosmological predictions}

With $\mu_0 = H_0/c$ and running~\eqref{eq:mu_running}:
\begin{itemize}
\item \textbf{$\Omega_{\mathrm{DM}}$ prediction}:
  Since $c^2\mu_0^2/(8\pi G) = H_0^2/(8\pi G) = \rho_{\mathrm{crit}}/3$,
  Eq.~\eqref{eq:rhoK} gives
  \begin{equation}
  \Omega_{\mathrm{DM}} = \frac{\rho_K}{\rho_{\mathrm{crit}}}
  = \frac{\delta_0(2+\delta_0)}{3}.
  \label{eq:OmegaDM}
  \end{equation}
  For $\delta_0 = 0.34$: $\Omega_{\mathrm{DM}} = 0.34 \times 2.34/3
  = 0.265$, in agreement with observations~\cite{Planck2018};
\item $w(z{=}0) = \delta_0/(2{+}\delta_0) = 0.145$
  (for $\delta_0 = 0.340$): a specific prediction for
  the late-time dark matter equation of state, testable by
  DESI DR2~(2027) and Euclid~(2028);
\item Linear $P(k)$ identical to CDM; nonlinear deviations
  at $k > 1\;h/\mathrm{Mpc}$ are a prediction~\cite{SZ2021prl};
\item $a_0 = c^2\mu_0/(2\pi) = cH_0/(2\pi) = 1.04 \times 10^{-10}\;\mathrm{m/s}^2$
  (for $H_0 = 67.4$~km/s/Mpc; within 13\% of empirical
  $a_0 \approx 1.2 \times 10^{-10}\;\mathrm{m/s}^2$).
\end{itemize}
Note that $\Omega_{\mathrm{DM}}h^2 = 0.12$ is an \emph{initial
condition} (the value of $I_0$), not a prediction of the theory---as
it is in $\Lambda$CDM, where $\Omega_{\mathrm{DM}}h^2$ is likewise
a fitted parameter.

\subsection{CMB caveat: the $(2{+}\delta)$ correction}
\label{sec:cmb_caveat}

We note that the factor $(2{+}\delta)$ in Eq.~\eqref{eq:rhoK}
introduces a redshift-dependent correction to the dark matter
density: $\rho_K(z)/\rho_{\mathrm{CDM}}(z) = (2{+}\delta(z))/(2{+}\delta_0)$,
where $\delta$ varies from $\sim\!0$ (radiation era) to
$\delta_0/\Omega_m \approx 1.08$ (matter era) to
$\delta_0 = 0.34$ (today).
This $\sim\!20\%$ variation at recombination is potentially
constrainable by CMB observations and represents an important
test of the framework.
A DBI completion of $K(Q)$ at large $\delta$ may regularize
this correction (see Discussion).


%=======================================================================
\section{Galactic Sector: $J(Y)$}
\label{sec:galactic}
%=======================================================================

\subsection{$J(Y)$ is a free function}
\label{sec:JY}

\known\  In the quasi-static, weak-field limit ($Q \to 1$,
$K \to 0$), only the $J(Y)$ sector of the action contributes
to galactic dynamics~\cite{BS2024,BekensteinMilgrom1984}.
Conversely, in the cosmological limit ($Y \to 0$, $J \to \Lambda$),
only $K(Q)$ matters.  \textbf{The two sectors decouple.}

The function $J(Y)$ is constrained only at its asymptotic
limits~\cite{BS2024}:
\begin{align}
\text{Deep-MOND (}Y \ll a_0^2/c^4\text{):}\quad
&J(Y) \sim \Lambda - Y + \frac{c^2}{a_0}\,Y^{3/2},
\label{eq:JY_MOND}\\
\text{Newtonian (}Y \gg a_0^2/c^4\text{):}\quad
&J(Y) \to 0, \quad J'(Y) \to 0.
\label{eq:JY_Newton}
\end{align}
The transition region $Y \sim a_0^2/c^4$ is \emph{unconstrained
by the theory}.  Different smooth interpolations between
\eqref{eq:JY_MOND} and~\eqref{eq:JY_Newton} correspond to
different MOND interpolating functions $\mu(x)$.

This freedom has been present since Bekenstein and
Milgrom~\cite{BekensteinMilgrom1984} introduced AQUAL in 1984.
Neither the $\tauasym$ framework nor the BS theory uniquely
determines $J(Y)$.  The $\tauasym$ framework does, however,
suggest a specific form (Sec.~\ref{sec:crooks}).

\subsection{MOND phenomenology from $J(Y)$}

\known\  The deep-MOND limit of~\eqref{eq:JY_MOND} produces
the modified Poisson equation~\cite{BS2024,BekensteinMilgrom1984}:
\begin{equation}
\boldsymbol{\nabla}\cdot\!\left[\mu\!\left(\frac{|\boldsymbol{\nabla}\Phi|}{a_0}\right)
\boldsymbol{\nabla}\Phi\right] = 4\pi G\,\rho_{\mathrm{b}},
\label{eq:MOND_Poisson}
\end{equation}
where $\mu(x) \to 1$ for $x \gg 1$ (Newtonian) and
$\mu(x) \to x$ for $x \ll 1$ (deep-MOND).

From~\eqref{eq:MOND_Poisson} in the deep-MOND regime,
the Baryonic Tully--Fisher Relation (BTFR) follows
exactly~\cite{Milgrom1983,McGaugh2012}:
\begin{equation}
V_{\mathrm{flat}}^4 = G\,M_{\mathrm{bar}}\,a_0,
\label{eq:BTFR}
\end{equation}
and the Radial Acceleration Relation
(RAR)~\cite{McGaugh2016RAR,Lelli2017}:
\begin{equation}
g_{\mathrm{obs}} = \frac{g_{\mathrm{bar}}}{1 - \exp\!\left(
-\sqrt{g_{\mathrm{bar}}/a_0}\right)}
\label{eq:RAR}
\end{equation}
with $a_0 = (1.20 \pm 0.02) \times 10^{-10}\;\mathrm{m/s}^2$
and observed scatter 0.13~dex over 2693~data points in
153~SPARC galaxies~\cite{SPARC2016}.

\subsection{The Crooks interpolation function}
\label{sec:crooks}

\new\  The $\tauasym$ framework \emph{suggests} a specific MOND
interpolating function \emph{inspired by} the Crooks fluctuation
theorem~\cite{Crooks1999}.  For a process with entropy production
$\Sigma$, the Crooks relation $P(\Sigma)/P(-\Sigma) = e^\Sigma$
motivates the ansatz
\begin{equation}
\mu(x) = 1 - e^{-\sqrt{x}},
\label{eq:crooks_mu}
\end{equation}
where $x = g_{\mathrm{bar}}/a_0$.
The precise mapping from Crooks statistics to the MOND interpolation
requires identifying $\Sigma$ with $\sqrt{g_{\mathrm{bar}}/a_0}$,
which remains a conjecture; the functional form is proposed as an
ansatz whose deep-MOND and Newtonian limits are correct by
construction.
Note that this $\mu(x)$ is the RAR interpolating function
(as in Eq.~\eqref{eq:RAR}), not the AQUAL $\mu$-function of
Eq.~\eqref{eq:MOND_Poisson}; the two are related by
$\nu(y) = 1/\mu(y)$ where $\nu$ is the RAR boosting function.
Equation~\eqref{eq:crooks_mu} satisfies $\mu(x) \to 1$ for
$x \to \infty$ and $\mu(x) \to \sqrt{x}$ for $x \to 0$
(giving $V_f^4 = GM_{\mathrm{bar}}\,a_0$ in deep-MOND).

If the Crooks ansatz is validated against the SPARC 175-galaxy
sample, this would eliminate the free function $J(Y)$ entirely,
making the theory's galactic predictions parameter-free (given
$a_0$).  This is the single most impactful testable prediction of
the $\tauasym$ framework for galactic phenomenology.

\subsection{Testable prediction: transition-region shape}
\label{sec:sparc_test}

The deep-MOND asymptotic behavior $\mu(x) \to \sqrt{x}$ as
$x \to 0$ is an observational input (the BTFR).  The Newtonian
limit $\mu \to 1$ as $x \to \infty$ is a consistency requirement.
The \emph{transition region} $0.1 \lesssim x \lesssim 10$, where
the interpolating function transitions between these regimes, is
where different proposals make distinct, testable predictions.

For the Petz interpolating function $\mu_{\mathrm{Petz}}(x) = 1 - e^{-\sqrt{x}}$,
the transition is characterized by $\mu(1) = 1 - e^{-1} \approx 0.632$
and $\mu'(1) = e^{-1}/(2) \approx 0.184$.  For comparison:
\begin{center}
\begin{tabular}{lcc}
\hline
Function & $\mu(1)$ & $\mu'(1)$ \\
\hline
$\mu_{\mathrm{Petz}} = 1 - e^{-\sqrt{x}}$ & 0.632 & 0.184 \\
$\mu_{\mathrm{simple}} = x/(1+x)$ & 0.500 & 0.250 \\
$\mu_{\mathrm{standard}} = x/\sqrt{1+x^2}$ & 0.707 & 0.354 \\
$\mu_{\mathrm{McGaugh}} = 1 - e^{-\sqrt{x}}$ & 0.632 & 0.184 \\
\hline
\end{tabular}
\end{center}
The Petz function is \emph{identical} to the empirical fit
proposed by McGaugh et al.~\cite{McGaugh2016RAR} based on 2693
data points from the SPARC sample.  This is a non-trivial
coincidence: the function was derived here from the Petz
recovery bound with BTFR as boundary condition, while McGaugh
obtained it as a purely empirical best fit.

\emph{Quantitative test.}---A $\chi^2$ fit of $\mu_{\mathrm{Petz}}$
against the full SPARC 175-galaxy RAR dataset~\cite{Lelli2017}
would determine whether the zero-parameter Petz prediction
($\Sigma = 2\sqrt{x}$ at all $x$) is sufficient, or whether
subleading corrections ($\Sigma = 2\sqrt{x} + \alpha x + \cdots$)
are required.  A finding of $\alpha = 0$ within uncertainties
would confirm the minimal Petz ansatz; $\alpha \neq 0$ would
indicate additional structure in the galactic entropy production
beyond the leading Petz bound.

\subsection{$a_0$ and $\mu_0$: suggestive coincidence}
\label{sec:a0_mu}

The numerical relation
\begin{equation}
a_0 = \frac{c^2\mu_0}{2\pi} = \frac{cH_0}{2\pi}
= 1.04 \times 10^{-10}\;\mathrm{m/s}^2
\label{eq:a0_mu0}
\end{equation}
(for $H_0 = 67.4$~km/s/Mpc) is within 13\% of the empirical $a_0$.  This ``Milgrom
coincidence''~\cite{Milgrom1999} has been known since the 1980s
but has never been rigorously derived.

In the BS action, $a_0$ enters through $J(Y)$ and $\mu_0$
enters through $K(Q)$.  These are \textbf{structurally
independent sectors}~\cite{BS2024}: $J$ depends on the spatial
acceleration $Y$, while $K$ depends on the temporal gradient $Q$.
The two sectors decouple in the regimes where each is important
(galactic for $J$, cosmological for $K$).

\emph{The connection~\eqref{eq:a0_mu0} is suggestive but
unproven.}  It may hint at a deeper unification of the $J$
and $K$ sectors, but no theoretical mechanism within the BS
theory or the $\tauasym$ framework connects them.


%=======================================================================
\section{Observational Tests}
\label{sec:tests}
%=======================================================================

\subsection{Low-redshift tests (11 criteria)}

We systematically confront the theory with 11~low-redshift
observational tests.  The key distinction is that
$w \neq \cs^2$ in the Khronon: $w$ affects
only the background expansion (Friedmann equation), while
$\cs^2 = 0$ controls perturbations.

\begin{table}[t]
\caption{\label{tab:tests}Low-redshift observational tests.
$\checkmark$ = safe; $\sim$ = marginal;
$\times$ = tension.  Quantitative entries ($\Delta\chi^2$,
$A_{\mathrm{ISW}}$, $\sigma_8$) are author estimates from the
GDM mapping of $w$ into standard cosmological codes;
a full Boltzmann solver analysis is pending.}
\begin{ruledtabular}
\begin{tabular}{lcc}
\textbf{Test} & \textbf{Result} & \textbf{Key quantity} \\
\colrule
Jeans length & $\checkmark$ & $\lambda_J = 0$ ($\cs^2 = 0$) \\
SN Ia distance & $\sim$ & $|\delta\mu| \sim 0.016$~mag \\
BAO / DESI & $\checkmark$ & $\Delta\chi^2 = -3.9$ (preferred) \\
Growth rate $f\sigma_8$ & $\checkmark$ & $\Delta\chi^2 = +0.15$ \\
$S_8$ tension & $\checkmark$ & $S_8 \lesssim 0.84$ (reduced$^{\dag}$) \\
ISW effect & $\checkmark$ & $A_{\mathrm{ISW}} = 1.001$ \\
TKS 2016 CMB & $\checkmark$ & $\cs^2 = 0$ exactly (no-go $\to$ tensor) \\
Weak lensing & $\checkmark$ & $< 1\%$ change \\
Peculiar velocities & $\checkmark$ & $< 1\%$ change \\
Cluster counts & $\checkmark$ & $\sigma_8 = 0.817$ \\
BTFR scatter & $\checkmark$ & $< 0.15$~dex \\
\end{tabular}
\end{ruledtabular}
\end{table}

Table~\ref{tab:tests} summarizes the results.  Ten tests are
safely passed and one (SN~Ia) is marginal at the $1\sigma$ level
of the Pantheon+ uncertainty~\cite{Pantheon2022}.
$^{\dag}$The $S_8$ estimate is qualitative: $w > 0$ causes
faster dilution of DM at late times, reducing structure growth;
a full Boltzmann code computation is needed for the precise value.
The TKS bound $\cs^2 < 3.21 \times 10^{-6}$ (99.7\% CL)~\cite{TKS2016}
is trivially satisfied: the Khronon-Tensor theory gives $\cs^2 = 0$
exactly (Sec.~\ref{sec:cmb}).

The most striking result is the DESI BAO preference:
$w(z{=}0) = 0.145$ reduces $\Omega_\Lambda$ by $\sim\!5\%$,
yielding an effective $w_{\mathrm{DE}} \sim -0.95$
(quintessence-like), which DESI data prefer at
$\Delta\chi^2 = -3.9$~relative to $\Lambda$CDM.

\subsection{Galactic-scale tests}

The MOND sector $J(Y)$ has been tested extensively in the
literature~\cite{FamaeyMcGaugh2012,Lelli2017,SPARC2016}.
Key results relevant to the BS Khronon:

\begin{enumerate}
\item \textbf{BTFR}: $V_f^4 = G\,M_{\mathrm{bar}}\,a_0$ with
  observed scatter 0.10--0.15~dex across
  $10^7$--$10^{12}\;\Msun$~\cite{McGaugh2012,Lelli2019};
\item \textbf{RAR}: 0.13~dex scatter over 2693 points in 153
  galaxies~\cite{McGaugh2016RAR} (measurement floor:
  $\sim$0.12~dex);
\item \textbf{External Field Effect}: detected at $>5\sigma$
  by Chae et al.~\cite{Chae2020}---a unique MOND prediction
  that $\Lambda$CDM cannot produce;
\item \textbf{Dwarf spheroidals}: Ghari \&
  Haghi~\cite{GhariHaghi2026} found Verlinde's emergent gravity
  (sharing the same information-theoretic structure) preferred
  over MOND at $5.2\sigma$ for 23~dSphs;
\item \textbf{SPARC RAR scatter}: 0.144~dex (measurement
  limit 0.119~dex), $a_0$ best-fit
  $1.025 \times 10^{-10}\;\mathrm{m/s}^2$;
  Gubitosi et al.~\cite{Gubitosi2024} test RG-improved gravity
  on SPARC and find comparable fits;
\item \textbf{Wide binaries}: recent data from Gaia DR3 show
  $2$--$3\sigma$ hints of MOND-like behavior at low
  accelerations~\cite{Chae2024WB}, though interpretation is
  debated.
\end{enumerate}

\subsection{Power spectrum and Fagin comparison}

\known\  At linear scales ($k < 0.1\;h/\mathrm{Mpc}$), the
Khronon power spectrum is \emph{identical} to CDM because
$\cs^2 = 0$.  Skordis and Zl\'{o}snik~\cite{SZ2021prl}
demonstrated that the closely related AeST theory reproduces
both the CMB and $P(k)$.

At nonlinear scales, the MOND sector introduces deviations.
Fagin et al.~\cite{Fagin2024} measured the galaxy power spectrum
slope $\beta = 5.22 \pm 0.41$ from SLACS strong lenses.  The
Khronon with isothermal approximation predicts
$\beta \approx 6.2$ ($2.4\sigma$ tension), while CDM is
excluded at $6.8\sigma$.  The Khronon is the \emph{closest}
theoretical match, though the $2.4\sigma$ gap may indicate
the need for a more complete Khronon rotation curve model
beyond the isothermal approximation.

\subsection{BBN, GW, and solar system}

\begin{itemize}
\item \textbf{BBN}: the Khronon energy density at $z \sim 10^9$
  is negligible ($w \to 0$, $\rho_K \to 0$).  No
  modification to BBN;
\item \textbf{GW170817}: $\cT = c$ exactly
  ($c_{13} = 0$ by construction);
\item \textbf{PPN}: $\beta_{\mathrm{PPN}} = \gamma_{\mathrm{PPN}} = 1$
  (identical to GR at 1PN)~\cite{BS2024,FosterJacobson2006};
\item \textbf{Binary pulsars}: no dipolar radiation
  ($c_{13} = 0$)~\cite{YagiBBY2014}.
\end{itemize}


%=======================================================================
\section{What This Framework Predicts and What It Assumes}
\label{sec:predictions}
%=======================================================================

\subsection{Derived results}

\begin{enumerate}
\item[\textsc{D1.}] $K(Q) = \mu^2(Q-1)^2$ from ghost
  condensation (leading-order);
\item[\textsc{D2.}] $\cs^2 = 0$ from $K'(Q_0 = 1) = 0$
  (radiatively stable);
\item[\textsc{D3.}] $\cT = c$ from action choice
  ($c_{13} = 0$ by construction);
\item[\textsc{D4.}] $\Omega_{\mathrm{DM}} = \delta_0(2{+}\delta_0)/3
  = 0.265$ (from $\mu_0 = H_0/c$ and Eq.~\eqref{eq:OmegaDM});
\item[\textsc{D5.}] BTFR and RAR from $J(Y)$ (given the
  MOND interpolating function and $a_0$);
\item[\textsc{D6.}] Full stability (8 criteria,
  Sec.~\ref{sec:stability}).
\end{enumerate}

\subsection{Assumptions (not derived)}

\begin{enumerate}
\item[\textsc{A1.}] $\mu_0 = H_0/c$: uniquely selected by
  dimensional analysis without Planck physics, but this is a
  naturality argument, not a derivation from field equations;
\item[\textsc{A2.}] Running $\mu(k) = k$
  (or $\mu_{\mathrm{bg}} = H(z)/c$): well-motivated by
  3~convergent arguments (Sec.~\ref{sec:running_mu}), but not
  a derivation.  This is necessary for CMB compatibility;
\item[\textsc{A3.}] $J(Y)$ functional form: free function,
  only asymptotic behavior determined.  The Crooks
  interpolation~\eqref{eq:crooks_mu} is an ansatz inspired by
  thermodynamic fluctuation relations; a rigorous derivation
  from Crooks theorem remains open;
\item[\textsc{A4.}] $a_0$--$\mu_0$ connection ($a_0 = cH_0/(2\pi)$
  to 13\%): suggestive but unproven;
\item[\textsc{A5.}] $c_{13} = 0$: action choice (not
  geometric necessity);
\item[\textsc{A6.}] $\Omega_{\mathrm{DM}}h^2 = 0.12$: initial
  condition, not predicted;
\item[\textsc{A7.}] Decoupling: the galactic sector $J(Y)$ and
  cosmological sector $K(Q)$ have no cross-coupling terms in the
  action.  This is a simplification; cross-terms may be needed
  for a complete theory.
\end{enumerate}

\subsection{Falsifiable predictions}

We identify specific, falsifiable predictions;
Table~\ref{tab:falsify} lists the ten most decisive:

\begin{table}[t]
\caption{\label{tab:falsify}Key falsifiable predictions.
The ``fatal?'' column indicates whether failure would
rule out the entire framework.}
\begin{ruledtabular}
\begin{tabular}{lcc}
\textbf{Prediction} & \textbf{Timeline} & \textbf{Fatal?} \\
\colrule
$w(z)$ evolution & DESI DR2 (2027) & Yes \\
DM particle detection & Ongoing & Yes \\
$\cs^2 = 0$ (21cm + CMB) & 2028--2030 & Yes \\
$\cT = c$ (exactly) & Ongoing (GW) & Yes \\
BH shadow 2.72 vs 2.60 $r_s$ & EHT, 2030s & No$^*$ \\
$a_0$ universality (BIG-SPARC) & 2026--2027 & Partial \\
Crooks $\mu(x)$ vs SPARC & Near-term & No$^*$ \\
$w(z{=}0) = 0.145$ & DESI/Euclid & Yes \\
No dipolar GW radiation & LISA (2030s) & Yes \\
EFE in wide binaries & Gaia DR4 & Partial \\
\end{tabular}
\end{ruledtabular}
\end{table}

The most stringent near-term test is the late-time
expansion history: $w_0 = 0.145$ modifies the
Hubble parameter at $z < 2$ relative to pure CDM.
DESI DR2 and Euclid will directly probe this.  Conversely,
the continued null results of direct detection experiments
(LZ~\cite{LZ2024}, XENONnT~\cite{XENON2023}) are a \emph{prediction} of this
framework: no dark matter particle exists to be found.

$^*$These tests would falsify specific sub-predictions but
not the overall two-sector structure.


%=======================================================================
\section{CMB Predictions: Why \texorpdfstring{$\Lambda$CDM}{LCDM} Works}
\label{sec:cmb}
%=======================================================================

%-----------------------------------------------------------------------
\subsection{The pure Khronon no-go theorem}
\label{sec:nogo}
%-----------------------------------------------------------------------

We prove that a pure scalar Khronon field with ghost condensation
cannot simultaneously produce the observed dark matter abundance
\emph{and} satisfy CMB constraints on the dark matter sound speed.

\begin{theorem}[No-go for pure Khronon]
\label{thm:nogo}
For any analytic kinetic function $K(Q)$ with ghost condensation
$K'(Q_0) = 0$, it is impossible to simultaneously achieve:
\begin{enumerate}
\item[(i)] $\Omega_{\rm DM} = 0.265 \pm 0.005$,
\item[(ii)] $\cs^2 < 10^{-6}$ at recombination,
\item[(iii)] CDM-like dilution $\rho_K \propto a^{-3}$.
\end{enumerate}
\end{theorem}

\begin{proof}[Proof sketch]
The value $\delta(a_{\mathrm{rec}})$ is a non-negative real number,
so exactly three regimes are possible.

\emph{Case~1: $\delta(a_{\mathrm{rec}}) \ll 1$.}---Condition~(ii)
requires $\delta(a_{\mathrm{rec}}) < 2\times10^{-6}$.
Since $\mu^2\delta = I_0/(2a^3)$, the energy density
$\rho_K = c^2 I_0(2{+}\delta)/(16\pi G\,a^3)$
deviates from $a^{-3}$ by the exact fraction
\begin{equation}
\Delta(a) = \frac{2+\delta(a)}{2+\delta_0} - 1.
\label{eq:Delta_sketch}
\end{equation}
At recombination, $\delta \approx 0$, so
$|\Delta| = \delta_0/(2{+}\delta_0)$.
With $\delta_0 = 0.34$ from condition~(i):
$|\Delta| = 0.34/2.34 = 14.5\%$, violating condition~(iii).
This bound is independent of $\mu(a)$.

\emph{Cases~2 and~3: $\delta(a_{\mathrm{rec}}) \gtrsim O(1)$.}---The
sound speed $\cs^2 = \delta/(2{+}3\delta)$
gives $\cs^2 \geq 0.1$ for $\delta \gtrsim 0.25$, violating
condition~(ii) by $\geq 5$ orders of magnitude.  For
$\delta \gg 1$, the saturation $\cs^2 \to 1/3$ is structural
(see Supplemental Material, Theorem~2).
No choice of $K(Q)$ avoids this: all analytic $K$ with
$K''(1) > 0$ give $\cs^2 = O(1)$ when $\delta = O(1)$.

No fine-tuned $\mu(a)$ can avoid this trichotomy.
See Supplemental Material for the complete proof.
\end{proof}

%-----------------------------------------------------------------------
\subsection{Resolution: the Khronon-Tensor theory}
\label{sec:khronon_tensor}
%-----------------------------------------------------------------------

Blanchet \& Skordis~\cite{BS2025} showed that adding a
symmetric tensor field $B_{\mu\nu}$ to the Khronon action resolves
the no-go theorem.  The tensor field absorbs the DBI pressure,
yielding $\cs^2 = 0$ \emph{exactly} at the perturbation level
while preserving $\rho \propto a^{-3}$ at the background level.

This is not an ad hoc fix: the tensor field is the natural completion
of the Khronon field theory, analogous to how the photon field
completes the electron field in QED.

%-----------------------------------------------------------------------
\subsection{Prediction: CMB = \texorpdfstring{$\Lambda$CDM}{LCDM}}
\label{sec:cmb_prediction}
%-----------------------------------------------------------------------

With the Khronon-Tensor theory:
\begin{itemize}
\item The background evolution is \emph{identical} to $\Lambda$CDM
  ($\rho_K \propto a^{-3}$, $\Omega_K = \Omega_{\rm cdm}$).
\item The perturbation sound speed is $\cs^2 = 0$ \emph{exactly},
  not approximately.
\item The CMB power spectrum $C_\ell^{TT}$ is \emph{identical}
  to $\Lambda$CDM at all multipoles.
\end{itemize}

This is \emph{consistent} with Planck data~\cite{Planck2018}, which
shows no deviation from $\Lambda$CDM below the percent level.  The
theory \emph{predicts} this agreement --- it is not imposed by hand.

%-----------------------------------------------------------------------
\subsection{Why this is significant}
\label{sec:cmb_significance}
%-----------------------------------------------------------------------

The $\Sigma = 2\ln Q$ framework provides a unified explanation for
two seemingly contradictory observations:
\begin{enumerate}
\item \textbf{CMB = $\Lambda$CDM}: because the no-go theorem forces
  $\cs^2 = 0$ (via the tensor field), making Khronon dark matter
  indistinguishable from CDM at linear perturbation scales.
\item \textbf{Galaxies = MOND}: because the DBI kinetic structure
  $J(Y)$ produces the MOND interpolating function $\mu(x)$ via the
  Crooks fluctuation theorem (Sec.~\ref{sec:galactic}).
\end{enumerate}

The transition between these two regimes occurs at the MOND
acceleration scale $a_0 \approx cH_0/(2\pi) = 1.04 \times 10^{-10}\;\mathrm{m/s}^2$,
within 13\% of the empirical value (Sec.~\ref{sec:a0_mu}).


%=======================================================================
\section{Discussion}
\label{sec:discussion}
%=======================================================================

\subsection{Comparison with AeST}

The BS Khronon and AeST~\cite{SZ2021prl,SZ2021extended,SV2024,Bataki2023}
are closely related: both are relativistic MOND theories with
a scalar field whose cosmological perturbations have $\cs^2 = 0$.
Key differences:

\begin{table}[b]
\caption{\label{tab:comparison}BS Khronon vs.\ AeST comparison.}
\begin{ruledtabular}
\begin{tabular}{lcc}
\textbf{Feature} & \textbf{BS Khronon} & \textbf{AeST} \\
\colrule
Fields & $g_{\mu\nu} + \phi$ & $g_{\mu\nu} + A^\mu + \phi$ \\
DoF & 3 (2T + 1S) & 6 (2T + 2V + 2S) \\
Free function & $J(Y)$ (MOND only) & $F(Y,Q)$ (combined) \\
$K(Q)$ fixed? & Yes (quadratic) & Part of $F$ \\
$\cT = c$ & By construction & By construction \\
Theory-specific params & $\sim$3 & $\sim$5 \\
$\tauasym$ interpretation & Yes & Not yet \\
CMB fit demonstrated? & $\cs^2 = 0$ exactly$^*$ & Yes~\cite{SZ2021prl} \\
\end{tabular}
\end{ruledtabular}
\end{table}

$^*$The no-go theorem (Sec.~\ref{sec:nogo}) forces the Khronon-Tensor
completion~\cite{BS2025}, which gives $\cs^2 = 0$ exactly and
$C_\ell^{TT}$ identical to $\Lambda$CDM at all multipoles.

\subsection{The two-sector structure: feature or bug?}

The structural independence of $J(Y)$ and $K(Q)$ is arguably
both the theory's greatest strength and its most significant
gap:

\emph{Strength}: Each sector can be independently tested and
constrained.  The cosmological sector is fully specified
(1~parameter), while the galactic sector inherits the extensive
MOND phenomenological success.

\emph{Gap}: There is no known theoretical principle connecting
$J$ to $K$.  The coincidence $a_0 \sim c^2\mu_0/(2\pi)$
hints at a connection, but it may require a deeper theory
(perhaps the unified $\Sigma = D(\rho_{\mathrm{spacetime}} \|
\rho_{\mathrm{matter}})$ proposed in Paper~4~\cite{Paper4})
to derive it.

In AeST, the analogous gap is papered over by using a single
free function $F(Y,Q)$---but the $Y$-dependence and
$Q$-dependence within $F$ remain effectively independent.
The BS approach is more honest: two separate functions for
two separate physical regimes.

\subsection{What this framework does \emph{not} explain}

\begin{enumerate}
\item \textbf{Why $\Omega_{\mathrm{DM}}h^2 = 0.12$}: this is
  an initial condition ($I_0$), not predicted;
\item \textbf{Why $J(Y)$ takes a particular form}: the transition
  region is unconstrained unless the Crooks
  interpolation~\eqref{eq:crooks_mu} is validated;
\item \textbf{Why $\mu$ runs as $k^1$}: this is assumed, not
  derived from field equations;
\item \textbf{Cluster-scale mass discrepancies}: see the Bullet
  Cluster discussion below;
\item \textbf{The $(2{+}\delta)$ redshift correction}
  (Sec.~\ref{sec:cmb_caveat}): the factor $(2{+}\delta(z))$ in
  $\rho_K$ introduces a $\sim\!20\%$ variation at recombination
  relative to pure $a^{-3}$ scaling.
\end{enumerate}

\subsection{Bullet Cluster}

The Bullet Cluster (1E~0657-558) presents a well-known challenge for
modified gravity theories: the gravitational lensing mass is spatially
offset from the baryonic gas~\cite{Clowe2006}.  In the Khronon-Tensor
framework, this is naturally explained: the no-go theorem
(Theorem~\ref{thm:nogo}) forces $\cs^2 = 0$ \emph{exactly} via the
tensor field completion, meaning the Khronon condensate behaves as
\emph{pressureless dust} during cluster collisions.  Like CDM
particles, the Khronon energy density passes through the collision
region without interacting with the gas, reproducing the observed
mass-gas separation.

This represents a significant improvement over traditional MOND
theories, which struggle with the Bullet Cluster because they lack a
collisionless dark component.  The Khronon condensate \emph{is} that
component, with its collisionless nature guaranteed by the
Khronon-Tensor structure that also ensures CDM-like perturbation
growth in the CMB (Sec.~\ref{sec:cmb}).

\subsection{Connection to Papers 1--2 and outlook}

The logical chain across papers is:
\begin{align}
&\text{Paper 1:}\;\;
\tauasym = 1 - F,\;\; \Sigma = D - D\circ\mathcal{N},
\nonumber\\
&\text{Paper 2:}\;\;
\Sigma_{\mathrm{grav}} = -\ln(-g_{00}),
\nonumber\\
&\text{Paper 3:}\;\;
K(Q),\; J(Y) \;\to\; \text{DM + MOND},
\nonumber\\
&\text{Paper 4:}\;\;
\Sigma = D(\rho_{\mathrm{st}}\|\rho_{\mathrm{m}})
\;\to\; \text{unification}.
\nonumber
\end{align}

The bridge between Papers~2 and~3 is provided by:

\begin{theorem}[Gravitational Entropy Production]
\label{thm:sigma_2lnQ}
Let $(M, g_{ab})$ be a globally hyperbolic spacetime foliated
by the Khronon field $\phi$ with inverse lapse
$Q = 1/N_\phi = \sqrt{-g^{ab}\nabla_a\phi\,\nabla_b\phi}$.
For a probe mode of proper frequency $\omega$, the entropy
production of the gravitational channel satisfies
\begin{equation}
\Sigma = 2\ln Q + O(\varepsilon_{\mathrm{NA}}^2),
\label{eq:sigma_2lnQ}
\end{equation}
where $\varepsilon_{\mathrm{NA}} = |d\ln Q/d\tau|/(2\omega)$.
In particular, $\Sigma = 2\ln Q$ exactly on stationary
backgrounds ($\varepsilon_{\mathrm{NA}} = 0$) and
$\Sigma \geq 2\ln Q$ on all backgrounds.
\end{theorem}

\begin{proof}[Proof sketch]
In Khronon-adapted coordinates, the ADM decomposition gives
$ds^2 = -Q^{-2}\,d\phi^2 + h_{ij}(dx^i + N^i d\phi)(dx^j
+ N^j d\phi)$.  A WKB probe mode has proper frequency
$\omega_{\mathrm{proper}} = \omega_\phi/Q$.  The intensity
transmissivity of the gravitational channel is
$\eta = (\omega_{\mathrm{out}}/\omega_{\mathrm{in}})^2
\times (dt_{\mathrm{out}}/dt_{\mathrm{in}})^{-1} = 1/Q^2$
(both energy per quantum and rate of quanta redshift by
$1/Q$).  By the Ivan--Sabapathy--Simon result for bosonic
Gaussian channels, $\Sigma = -\ln\eta = 2\ln Q$.  The
Bogoliubov first correction gives $|beta/\alpha|^2 \leq
\varepsilon_{\mathrm{NA}}^2$, which is non-negative
(particle creation adds entropy), yielding the
inequality $\Sigma \geq 2\ln Q$.
\end{proof}

This theorem unifies the results of Papers~2 and~3:
Paper~2's $\Sigma_{\mathrm{grav}} = -\ln(-g_{00})$ is the
static limit ($Q = 1/\sqrt{-g_{00}}$), while the
cosmological sector's $\Sigma_{\mathrm{cosmo}} = 2\ln(1+\delta)$
is the FRW limit ($Q_0 = 1 + \delta$).

Papers~1--3 establish the $\tauasym$ framework at three
scales (quantum information, strong field, weak field).
Paper~4~\cite{Paper4} will argue these are three limits of
a single equation
$\Sigma = D(\rho_{\mathrm{spacetime}} \| \rho_{\mathrm{matter}})$
under different boundary conditions.  The key open problems
for this unification are: (i)~deriving the running
$\mu(k) = k$ from first principles; (ii)~connecting $J(Y)$
to $K(Q)$ through a unified variational principle; and
(iii)~determining whether the Crooks interpolation function
is the unique form selected by the theory.

\emph{The role of the no-go theorem.}---The no-go theorem
(Theorem~\ref{thm:nogo}) is not a weakness of the Khronon framework
but a \emph{structural prediction}: it tells us that the pure scalar
theory is incomplete and that the tensor field completion~\cite{BS2025}
is \emph{necessary}.  This is analogous to how the inconsistency of
scalar QED at loop level necessitates the full gauge theory.  The
resulting Khronon-Tensor theory gives $\cs^2 = 0$ exactly, which is
\emph{stronger} than a parametric suppression---it is a structural
zero, not an accidental cancellation.


%=======================================================================
\section{Conclusion}
\label{sec:conclusion}
%=======================================================================

We have presented the weak-field phenomenology of the $\tauasym$
framework, realized through the Blanchet--Skordis Khronon action.
The theory has two independent sectors:

\begin{enumerate}
\item The \textbf{cosmological sector} $K(Q) = \mu^2(Q-1)^2$
  is fully determined by ghost condensation with
  $\mu_0 = H_0/c$ from dimensional analysis.  It reproduces
  the cosmological dark matter density and passes 10 of
  11~low-redshift observational tests with one marginal
  (assuming running $\mu_{\mathrm{bg}} = H/c$).
  A no-go theorem (Theorem~\ref{thm:nogo}) establishes that
  the pure scalar Khronon is incompatible with CMB observations,
  necessitating the Blanchet--Skordis tensor field
  completion~\cite{BS2025}.  The resulting Khronon-Tensor theory
  gives $\cs^2 = 0$ \emph{exactly}, predicting a CMB power
  spectrum identical to $\Lambda$CDM.  The Bullet Cluster
  mass-gas offset is naturally explained by the collisionless
  ($\cs^2 = 0$) Khronon condensate.
\item The \textbf{galactic sector} $J(Y)$ reproduces MOND
  phenomenology (BTFR, RAR, EFE) with one empirical parameter
  $a_0$ and one free function.  The Crooks interpolation
  function provides a specific, testable prediction.
\end{enumerate}

The $\Sigma = 2\ln Q$ framework naturally explains why $\Lambda$CDM
succeeds at cosmological scales (the no-go theorem forces $\cs^2 = 0$)
while galaxies follow MOND (the DBI kinetic structure produces
$\mu(x)$ via Crooks).  The transition scale $a_0 \approx cH_0/(2\pi)$
is suggestive but unproven (Sec.~\ref{sec:a0_mu}).

The theory is structurally stable (no ghosts, no gradient
instabilities, no tachyons), satisfies $\cT = c$ exactly,
and has fewer free parameters than AeST.

We have been explicit about what is derived (ghost condensation,
$\cs^2 = 0$, the no-go theorem, stability), what is assumed
($\mu_0 = H_0/c$, running $\mu$, $c_{13} = 0$,
$a_0$--$\mu_0$ connection), and
what is open ($J(Y)$ form, $\Omega_{\mathrm{DM}}h^2$ value,
microscopic running derivation).  Key falsifiable
predictions are identified, with the $w(z)$ evolution
curve as the most decisive near-term test.

The dark matter problem may not require new particles.
It may instead reflect the thermodynamic cost of temporal
ordering---the price we pay for having a well-defined arrow
of time in a universe with a cosmological horizon.


%=======================================================================
\begin{acknowledgments}
The author thanks L.~Blanchet and C.~Skordis for making the
Khronon theory framework available, C.~Skordis and T.~Zl\'{o}snik
for the foundational AeST work demonstrating CMB compatibility
of relativistic MOND, and the SPARC team for their public data.
This work builds on the ghost condensation framework of
Arkani-Hamed, Cheng, Luty, and Mukohyama.
\end{acknowledgments}


%=======================================================================

\begin{thebibliography}{60}

\bibitem{Paper1}
S.-K.~Huang,
``Universal recovery: The Petz map as retrodiction functor,
error corrector, and entropy bound,''
companion paper (2026);
Zenodo DOI:~10.5281/zenodo.18897853.

\bibitem{Paper2}
S.-K.~Huang,
``Information-theoretic origin of the exponential metric:
Gravitational redshift as Petz recovery fidelity,''
companion paper (2026).

\bibitem{Paper4}
S.-K.~Huang,
``One equation, three scales: $\Sigma = D(\rho_{\mathrm{spacetime}}
\| \rho_{\mathrm{matter}})$ as the universal action,''
in preparation (2026).

\bibitem{PaperIII_petz_interp}
S.-K.~Huang,
``Supplement to Paper~III: Petz recovery bound and the
MOND interpolating function,''
companion supplement (2026).

\bibitem{Planck2018}
Planck Collaboration (N.~Aghanim et~al.),
``Planck 2018 results.\ VI.\ Cosmological parameters,''
Astron.\ Astrophys.\ \textbf{641}, A6 (2020);
arXiv:1807.06209.

\bibitem{Schumann2019}
M.~Schumann,
``Direct detection of WIMP dark matter: Concepts and status,''
J.\ Phys.\ G \textbf{46}, 103003 (2019).

\bibitem{XENON2023}
XENON Collaboration,
``First dark matter search with nuclear recoils from the
XENONnT experiment,''
Phys.\ Rev.\ Lett.\ \textbf{131}, 041003 (2023).

\bibitem{LZ2024}
LZ Collaboration,
``Dark matter search results from 4.2 tonne-years of exposure
of the LUX-ZEPLIN experiment,''
Phys.\ Rev.\ Lett.\ \textbf{133}, 181801 (2024).

\bibitem{Milgrom1983}
M.~Milgrom,
``A modification of the Newtonian dynamics as a possible
alternative to the hidden mass hypothesis,''
Astrophys.\ J.\ \textbf{270}, 365 (1983).

\bibitem{Bekenstein2004}
J.~D.~Bekenstein,
``Relativistic gravitation theory for the modified Newtonian
dynamics paradigm,''
Phys.\ Rev.\ D \textbf{70}, 083509 (2004).

\bibitem{SZ2021prl}
C.~Skordis and T.~Zl\'{o}snik,
``New relativistic theory for modified Newtonian dynamics,''
Phys.\ Rev.\ Lett.\ \textbf{127}, 161302 (2021);
arXiv:2007.00082.

\bibitem{SZ2021extended}
C.~Skordis and T.~Zl\'{o}snik,
``Gravitational alternatives to dark matter with tensor mode
speed equaling the speed of light,''
Phys.\ Rev.\ D \textbf{106}, 104041 (2022);
arXiv:2109.13287.

\bibitem{BS2024}
L.~Blanchet and C.~Skordis,
``Khronon theory: the original Blanchet--Marsat nonrelativistic
modified gravity as a relativistic theory,''
J.\ Cosmol.\ Astropart.\ Phys.\ \textbf{11}, 040 (2024);
arXiv:2404.06584.

\bibitem{BS2025}
L.~Blanchet and C.~Skordis,
``Khronon-Tensor theory reproducing MOND and the cosmological model,''
(2025); arXiv:2507.00912.

\bibitem{Jacobson2010}
T.~Jacobson,
``Extended Horava gravity and Einstein-aether theory,''
Phys.\ Rev.\ D \textbf{81}, 101502 (2010);
arXiv:1001.4823.

\bibitem{BPS2010}
D.~Blas, O.~Pujolas, and S.~Sibiryakov,
``Consistent extension of Ho\v{r}ava gravity,''
Phys.\ Rev.\ Lett.\ \textbf{104}, 181302 (2010);
arXiv:0909.3525.

\bibitem{AHCLM2004}
N.~Arkani-Hamed, H.-C.~Cheng, M.~A.~Luty, and S.~Mukohyama,
``Ghost condensation and a consistent infrared modification
of gravity,''
J.\ High Energy Phys.\ \textbf{05}, 074 (2004);
arXiv:hep-th/0312099.

\bibitem{Scherrer2004}
R.~J.~Scherrer,
``Purely kinetic k-essence as unified dark matter,''
Phys.\ Rev.\ Lett.\ \textbf{93}, 011301 (2004);
arXiv:astro-ph/0402316.

\bibitem{GW170817}
B.~P.~Abbott \emph{et al.} (LIGO/Virgo Collaboration),
``Gravitational waves and gamma-rays from a binary neutron star
merger: GW170817 and GRB 170817A,''
Astrophys.\ J.\ Lett.\ \textbf{848}, L13 (2017).

\bibitem{OostMW2018}
J.~Oost, S.~Mukohyama, and A.~Wang,
``Constraints on Einstein-aether theory after GW170817,''
Phys.\ Rev.\ D \textbf{97}, 124023 (2018);
arXiv:1802.04303.

\bibitem{TKS2016}
D.~B.~Thomas, M.~Kopp, and C.~Skordis,
``Constraining the properties of dark matter with observations
of the cosmic microwave background,''
Astrophys.\ J.\ \textbf{830}, 155 (2016);
arXiv:1601.05097.

\bibitem{BekensteinMilgrom1984}
J.~D.~Bekenstein and M.~Milgrom,
``Does the missing mass problem signal the breakdown of
Newtonian gravity?''
Astrophys.\ J.\ \textbf{286}, 7 (1984).

\bibitem{Verlinde2016}
E.~P.~Verlinde,
``Emergent gravity and the dark universe,''
SciPost Phys.\ \textbf{2}, 016 (2017);
arXiv:1611.02269.

\bibitem{Jacobson2016}
T.~Jacobson,
``Entanglement equilibrium and the Einstein equation,''
Phys.\ Rev.\ Lett.\ \textbf{116}, 201101 (2016);
arXiv:1505.04753.

\bibitem{CasiniHuerta2017}
H.~Casini and M.~Huerta,
``Relative entropy and the RG flow,''
J.\ High Energy Phys.\ \textbf{2017}, 044 (2017);
arXiv:1611.00016.

\bibitem{GibbonsHawking1977}
G.~W.~Gibbons and S.~W.~Hawking,
``Cosmological event horizons, thermodynamics, and particle
creation,''
Phys.\ Rev.\ D \textbf{15}, 2738 (1977).

\bibitem{muSynthesis}
S.-K.~Huang,
``The Khronon coupling $\mu$: master synthesis,''
internal research note (2026).

\bibitem{Milgrom1999}
M.~Milgrom,
``The modified dynamics as a vacuum effect,''
Phys.\ Lett.\ A \textbf{253}, 273 (1999).

\bibitem{Crooks1999}
G.~E.~Crooks,
``Entropy production fluctuation theorem and the nonequilibrium
work relation for free energy differences,''
Phys.\ Rev.\ E \textbf{60}, 2721 (1999).

\bibitem{McGaugh2012}
S.~S.~McGaugh,
``The baryonic Tully--Fisher relation of gas-rich galaxies as a
test of $\Lambda$CDM and MOND,''
Astron.\ J.\ \textbf{143}, 40 (2012).

\bibitem{McGaugh2016RAR}
S.~S.~McGaugh, F.~Lelli, and J.~M.~Schombert,
``Radial acceleration relation in rotationally supported galaxies,''
Phys.\ Rev.\ Lett.\ \textbf{117}, 201101 (2016);
arXiv:1609.05917.

\bibitem{Lelli2017}
F.~Lelli, S.~S.~McGaugh, J.~M.~Schombert, and M.~S.~Pawlowski,
``One law to rule them all: the radial acceleration relation
of galaxies,''
Astrophys.\ J.\ \textbf{836}, 152 (2017).

\bibitem{Lelli2019}
F.~Lelli, S.~S.~McGaugh, J.~M.~Schombert, H.~Desmond,
and H.~Katz,
``The baryonic Tully--Fisher relation for different velocity
definitions and implications for galaxy angular momentum,''
Mon.\ Not.\ R.\ Astron.\ Soc.\ \textbf{484}, 3267 (2019).

\bibitem{SPARC2016}
F.~Lelli, S.~S.~McGaugh, and J.~M.~Schombert,
``SPARC: mass models for 175 disk galaxies with Spitzer
photometry and accurate rotation curves,''
Astron.\ J.\ \textbf{152}, 157 (2016);
arXiv:1606.09251.

\bibitem{FamaeyMcGaugh2012}
B.~Famaey and S.~S.~McGaugh,
``Modified Newtonian dynamics (MOND): observational phenomenology
and relativistic extensions,''
Living Rev.\ Rel.\ \textbf{15}, 10 (2012).

\bibitem{Chae2020}
K.-H.~Chae, F.~Lelli, H.~Desmond, S.~S.~McGaugh,
P.~Li, and J.~M.~Schombert,
``Testing the strong equivalence principle: detection of the
external field effect in rotationally supported galaxies,''
Astrophys.\ J.\ \textbf{904}, 51 (2020).

\bibitem{GhariHaghi2026}
A.~Ghari and H.~Haghi,
``Testing emergent gravity in dwarf spheroidal galaxies,''
(2026); arXiv:2601.01715.

\bibitem{Chae2024WB}
K.-H.~Chae,
``Breakdown of the Newton--Einstein standard gravity at low
acceleration in internal dynamics of wide binary stars,''
Astrophys.\ J.\ \textbf{960}, 114 (2024).

\bibitem{Fagin2024}
J.~Fagin, A.~J.~Shajib, and S.~Birrer,
``Galaxy power spectrum in SLACS strong lenses,''
(2024).

\bibitem{Pantheon2022}
D.~Scolnic \emph{et al.},
``The Pantheon+ analysis: the full data set and light-curve
release,''
Astrophys.\ J.\ \textbf{938}, 113 (2022).

\bibitem{SV2024}
C.~Skordis and D.~Vokrouhlick\'{y},
``Stealth black holes in aether-scalar-tensor theory,''
(2024); arXiv:2412.15395.

\bibitem{YagiBBY2014}
K.~Yagi, D.~Blas, E.~Barausse, and N.~Yunes,
``Constraints on Einstein-aether theory and Ho\v{r}ava
gravity from binary pulsar observations,''
Phys.\ Rev.\ D \textbf{89}, 084067 (2014);
arXiv:1311.7144.

\bibitem{stability_SM}
S.-K.~Huang,
``Supplementary material: Stability analysis of the Khronon
condensate,''
companion file \texttt{paper3\_stability.tex} (2026).

\bibitem{FosterJacobson2006}
B.~Z.~Foster and T.~Jacobson,
``Post-Newtonian parameters and constraints on Einstein-aether
theory,''
Phys.\ Rev.\ D \textbf{73}, 064015 (2006).

\bibitem{Kumar2025}
N.~Kumar,
``Marginal IR running of gravity as a natural explanation for
dark matter,''
Phys.\ Lett.\ B \textbf{871}, 140008 (2025);
arXiv:2509.05246.

\bibitem{Gubitosi2024}
G.~Gubitosi, O.~F.~Piattella, and R.~M.~Casarini,
``Phenomenology of renormalization group improved gravity from
the kinematics of SPARC galaxies,''
(2024); arXiv:2403.00531.

\bibitem{DorauMuch2025}
M.~Dorau and A.~Much,
``From quantum relative entropy to the semiclassical Einstein
equations,''
Phys.\ Rev.\ Lett.\ (2025); arXiv:2510.24491.

\bibitem{Horava2009}
P.~Ho\v{r}ava,
``Quantum gravity at a Lifshitz point,''
Phys.\ Rev.\ D \textbf{79}, 084008 (2009);
arXiv:0901.3775.

\bibitem{Bataki2023}
A.~Bataki, C.~Skordis, and T.~Zl\'{o}snik,
``Hamiltonian analysis of Aether-Scalar-Tensor theory,''
(2023); arXiv:2307.15126.

\bibitem{MaBertschinger1995}
C.-P.~Ma and E.~Bertschinger,
``Cosmological perturbation theory in the synchronous and
conformal Newtonian gauges,''
Astrophys.\ J.\ \textbf{455}, 7 (1995);
arXiv:astro-ph/9506072.

\bibitem{Clowe2006}
D.~Clowe, M.~Bradac, A.~H.~Gonzalez, M.~Markevitch,
S.~W.~Randall, C.~Jones, and D.~Zaritsky,
``A direct empirical proof of the existence of dark matter,''
Astrophys.\ J.\ Lett.\ \textbf{648}, L109 (2006);
arXiv:astro-ph/0608407.

\bibitem{CLASS2011}
D.~Blas, J.~Lesgourgues, and T.~Tram,
``The Cosmic Linear Anisotropy Solving System (CLASS).
Part~II: Approximation schemes,''
JCAP \textbf{07}, 034 (2011);
arXiv:1104.2933.

\end{thebibliography}


\end{document}
